Litrepl: Literate Paper Processor Promoting Transparency More Than Reproducibility

### 2024-12-28 paper/paper.tex
Update env; Update paper

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{svg}
\usepackage{natbib}

@@

@@

Concurrently, another concept of human-computer interaction called the
Read-Evaluate-Print Loop (REPL) gained traction, notably within the LISP
community and through the APL language designed for mathematical computations.
The combination of a command-line interface and a language interpreter enables
incremental and interactive programming, allowing users to directly modify the
interpreter state. By maintaining human involvement in the loop, this approach
facilitates human thought processes \cite{McCarthy1959recfun, Iverson1962apl}.

@@

6. First, we utilize simple bidirectional text streams for inter-process
communication with language interpreters to evaluate code. Second, we advocate
for the reuse of existing text document formats. In both the Markdown and LaTeX
formats that we have implemented, simplified parsers are utilized to distinguish
code and result sections from the rest of the document. As of now, we support
Python and Shell interpreter families, as well as a custom AI communication
interpreter. Finally, we strive to leverage operating system facilities as much
as possible.

@@

\begin{figure*}[t]
  \centering
  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
  \caption{Caption describing the image shown in the figure.}
  \label{fig:example}
\end{figure*}

\subsection{Inter-process communication}

When the file appears on the Litrepl's stdin stream, it uses Earley parser to
reliably recognize code and result sections.

@@

\section*{References}
\bibliographystyle{plain} % Set the bibliography style
\bibliography{paper} % Include the .bib file named 'paper'

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{svg}

\usepackage{biblatex}
\addbibresource{paper.bib}

@@

\newcommand{\ls}{\begin{itemize}\item}
\newcommand{\li}{\item}
\renewcommand{\le}{\end{itemize}}

\newcommand{\es}{\begin{enumerate}\item}
\newcommand{\ei}{\item}
\newcommand{\ee}{\end{enumerate}}

@@

Concurrently, another concept of human-computer interaction called the
Read-Evaluate-Print Loop (REPL) gained traction, notably within the LISP
community and through the APL language designed for mathematical computations.
The combination of a command-line interface and a language interpreter enables
incremental and interactive programming, allowing users to directly modify the
interpreter state. By maintaining human involvement in the loop, this approach
facilitates human thought processes \cite{Spence1975apl, McCarthy1959recfun,
Iverson1962apl}.

@@

6. First, we utilize simple bidirectional text streams for inter-process
communication with language interpreters to evaluate code. Second, we advocate
for the reuse of existing text document formats. In both the Markdown and LaTeX
formats that we have implemented, simplified parsers are utilized to distinguish
code and result sections from the rest of the document. As of now, we support
Python and Shell interpreter families, as well as a custom AI communication
interpreter. Finally, we strive to leverage POSIX \cite{POSIX2024} operating
system facilities as much as possible.

@@

Litrepl is implemented in Python as a command-line text processor utility. Its
primary function is to take a text document as input, process it according to
specified command line arguments and environment settings, and then output the
resulting document through its standard output.

The operation of Litrepl is best illustrated through the following example:

\begin{verbatim}
$ cat input.tex
\begin{python}
print("Hello")
\end{python}
\begin{result}
\end{result}

$ cat input.tex | litrepl
\begin{python}
print("Hello")
\end{python}
\begin{result}
Hello
\end{result}
\end{verbatim}

In this example, the input document \verb|input.tex| contains a Python code
block that outputs the string "Hello". When the document is piped through
Litrepl, it processes the code block and inserts the result directly into the
document where specified.

\begin{figure*}[t]
  \centering
  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
  \caption{Caption describing the image shown in the figure.}
  \label{fig:resource-allocation}
\end{figure*}

\subsection{Interfacing Interpreters}

Litrepl communicates with interpreters using two uni-directional text streams:
one for writing input and another for reading outputs. To establish effective
communication, the interpreter should conform to the following general
assumptions:

\begin{itemize}
  \item Synchronous single-user mode, which is implemented in most interpreters.
  \item A capability to disable command line prompts. Litrepl relies on the echo
        response, as described below, rather than on prompt detection.
  \item The presence of an echo command or equivalent. The interpreter must be
        able to echo an argument string provided by the user in response to the
        echo command.
\end{itemize}

In Litrepl, these details are hardcoded for several prominent interpreter
families, which we refer to as \textbf{interpreter classes}. At the time of
writing, Litrepl supports three classes: \textbf{python}, \textbf{sh}, and
\textbf{ai}. Using command line arguments, users can configure how to map code
section labels to the correct class and specify which interpreter command to
execute to start a session for each class.

1. In order to start interpreter sessions, Litrepl allocates resources as shown
on the Figure \ref{fig:resource-allocation}. First, the auxilary directory is
created using the name, computed from the current interpreter class, working
directory and the name of a system folder for temporary files.  Next, the two
POSIX pipes, in.pipe and out.pipe, are created. Finally, the configured
interpreter is started with the pipes attached.

@@

\section*{References}
\printbibliography
